---
title: "Your Introduction to Data with R"
author: "Ayjeren R. Bekmuratovna, Ingo Rohlfing"
output: learnr::tutorial
runtime: shiny_prerendered
description: This is a tutorial that introduces basics of data analysis using R.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd(rstudioapi::getActiveProject())
library(learnr)
```


## Topic 1: Introduction to R and RStudio
A video tutorial introducing R, RStudio, how to download them, main panels and basic logic: to be inserted. 

![]()

### Exercise 

*Here's a simple exercise with an empty code chunk. Multiple options are correct.* 

What can R be used for?:

```{r quiz_what_is_r, echo=FALSE}
question("R can be used for",
         answer("Basic calculation", correct = TRUE),
         answer("Data import and management", correct = TRUE),
         answer("Data creation"),
         answer("Data analysis", correct = TRUE),
         answer("Data visualisation", correct = TRUE),
         allow_retry = TRUE
)
```


## Topic 2: Grammar of R (Operators)
As any other programming language, R has its own unique grammar. Arguably, the 
most important part of R is its usage of *verbs*. There are two primary verb 
groups in R: 

* Operators
* Functions 

We first dive into operators and consider verbs later.

### Operators

Operators are basic verbs in R. They allow you to perform basic operations with 
data. They are akin to auxiliary verbs in English and can be used on their own 
or together with more complex functions. There are:

* Arithmetic operators (used to conduct arithmetic operations on data)
* Logical operators (used primarily for comparisons of data)
* Assignment operators (used to assign names to existing values)

We now discuss each in turn.

## Topic 2.1: Arithmetic operators
First, lets look at arithmetic operators. There are 6 primary arithmetic 
operators. Each of them allows to conduct simple arithmetic tasks such as adding or dividing numbers: 

1. Addition: `+`
2. Subtraction: `-`
3. Multiplication: `*`
4. Division: `/`
5. Exponentiation: `^`
6. Modulo: `%%`. Modulo divides two numbers and returns the remainder. For an 
example, press "Run code" below:

```{r modulo, exercise=TRUE, exercise.lines = 1}
7 %% 5
```


#### Exercises for arithmetic operators
Replace the alphabetic placeholders `a` and `b` below with numbers of your choice.

```{r a_operators, eval = F, exercise=TRUE, exercise.lines = 18}
# Division
a/b

# Multiplication
a*b

# Subtraction
a-b

# Addition
a+b

# Exponentiation
a^b

# Modulo
a %% b
```



## Topic 2.2: Logical operators

Logical operators are verbs that allow one to realize logical operations. There are 6 logical operators: 

* Less than: `<`
* Greater than: `>`
* Less than or equal to: `<=`
* Greater than or equal to: `>=`
* Equal: `==`
* Unequal: `!=`

R indicates the answer to a logical comparison by returning one of two answers:
`TRUE` or `FALSE`. `TRUE` indicates that the logical statement that you 
evaluated is true (correct). `FALSE` indicates that the evaluated statement is 
wrong (not correct). The following piece of code returns FALSE as an answer 
because 5 is **not** less than 4:

```{r less_than, exercise=TRUE, exercise.lines = 1}
5 < 4
```


#### Exercise for arithmetic operators
Replace the alphabetic placeholders `a` and `b` below with numbers of your
choice and exercise using logical operators.

```{r l_operators, eval = F, exercise=TRUE, exercise.lines = 18}
# Less than
a < b

# More than
a > b

# Less than or equal
a <= b

# More than or equal
a >= b

# Equal
a == b

# Not equal
a != b
```




## Topic 2.3: Assignment operators
An assignment operator are a special type of verb. It is are akin to the verb 
**to be**. There are two of them that work in almost the same way:

* `<-`
* `=`

Both are used to give (= assign to) a name to a value or a set of values. With 
the following code, we tell R that we want to save the number 1914 under the 
name `start_of_WWI` (WW = World War). This name alongside with the number will 
be stored in the R environment, which can be understood as the memory you are 
working with in a R session. Named entities are generally referred to as 
**objects** in R (more on this later). Once you have created an object, you can 
print it on the screen by typing `start_of_WWI` in the R console. 


The second assignment operator `<-` does exactly what `=` sign does. Here, we 
only assign a value to the object `start_of_WWI`. By default, you do not see any
output in the console when creating a new object. If the command works, nothing
happens in the console and you will only notice that a new object has been 
added to the R environment.

Here is an example. You can run the code and then type `start_of_WWI` in the 
chunk and execute this line to see that you created this object (put the cursor 
somewhere in the third line and hit Strg+Enter on your keyboard).

```{r start-WWI, exercise=TRUE, exercise.lines = 3}
start_of_WWI = 1914

start_of_WWI
```

We now create a second object with the other operator `<-`. The name of the 
object that you create always has to be on the side of the arrow's head. The 
second line of the chunk shows that one also work with `->`. The default use is 
the first line with `<-`.
```{r end-WWI, exercise=TRUE, exercise.lines = 3}
end_of_WWI <- 1918
1918 -> end_of_WWI
```

In practice, **one always uses the `<-` as an assignment operator in R** for
reasons we do not need to dig into here. For this reason, we will only use 
`<-` in the following. *Note*: Functions use  `=` 
internally instead of `<-`, for example when creating a variable inside a 
dataframe. The assignment operator is superior for creating objects *with*
a function.

With the objects `start_of_WWI` and `end_of_WWI`, you can now calculate the 
length of World War I *without* having to create a new object using an 
assignment operator (just as with an ordinary calculator). 

```{r WWI, include = F}
end_of_WWI <- 1918
start_of_WWI <- 1914

```

```{r assign_diff, exercise = TRUE, exercise.lines = 3, exercise.setup = "WWI"}
end_of_WWI-start_of_WWI
```

Now that you know how to create objects, here is some [good coding practice](https://style.tidyverse.org/syntax.html){target="blank"} to follow 
when naming objects. Usually, you should only use:

- lowercase letters;
- numbers;
- underscores (`_`, aka snakecase) 

*If* you want to use a blank space in an object name, you can do it by putting 
backticks (``) around the object name.

```{r object_w_blanks, exercise = TRUE, exercise.lines = 3, exercise.setup = "WWI"}
`duration of WWI` <- end_of_WWI-start_of_WWI

`duration of WWI` 
```


#### Exercise for assignment operators

In the code below,

1. assign your date of birth you the object `date_of_birth` (you can use a 
fictional year of birth, if you like).
2. Calculate your age in years by subtracting `date_of_birth` from the current
year we have.

```{r create_obj, exercise=TRUE, exercise.lines = 5}

```



## Topic 3: Grammar of R (Functions)
Functions are similar to verbs and perform an action. The action could be a 
calculation, the plotting of a graph etc. For example, `sum()` simply adds the 
numbers that you specify.

```{r sum_func, exercise=TRUE, exercise.lines = 3}
sum(1, 2)
```

The function adds any number of digits and returns the result of the summation 
action. In the script below, you can replace the alphabetic placeholders with 
actual numbers of your own choice:

```{r sum_func_1, eval = F, exercise=TRUE, exercise.lines = 3}
sum(a, b, c)
```

Every function performs a specific task. Functions always have at least one 
*argument*, very often more than one. `sum()` has at a minimum two arguments
because one needs two numbers for building a sum. In the chunks above, we 
represented the arguments by the letters `a`, `b` and `c`c. You can get an idea 
about the arguments of a function by typing `help()`, which is a function 
itself, and include the name of the function *without* brackets 
(e.g., `help(sum)`). When you check the help file for `sum()`, you will notice 
that the only argument are three dots (`...`). The 'Arguments' section of the 
help fule then states that they stand for "numeric or complex or logical vectors".

When you work with functions, you need to keep two things in mind. A function has:

- a prespecified order of arguments;
- default values for selected arguments (depends on the function)

Let's clarify this with the simple function for rounding decimal places: `round()`. 
It has two arguments. In this order:

- `x`: The number to be rounded
- `digits = 0`

Now play around with this function. What is the result when you run this chunk?

```{r default-digits, exercise = TRUE, exercise.lines = 1}
round(2.654)
```

You get a natural number because the default setting is zero digits. If you want
it rounded to one decimal place, you need this.

```{r two-digits, exercise = TRUE, exercise.lines = 1}
round(2.654, digits = 1)
```

You can also write it in these two equivalent ways. In the first line, R takes 
the numbers and uses them as input for the arguments as they are listed in the 
help file. The second line *reverses* the order of the arguments. Since we tell 
R that the first number is the number of digits, the second number, 2.654, is
automatically taken as the number to be rounded, giving us the same result as 
the first line.

```{r digits-first, exercise = TRUE, exercise.lines = 3}
round(2.654, 2)

round(digits = 2, 2.654)
```

It is particularly important to pay attention to the order of arguments and 
their default values when:

- you are not deeply familiar with a function;
- the number of arguments is large; 
- the function is regularly updated by a programmer because the types of 
arguments, their number and default values might have changed.


## Topic 4: Grammar of R (Data Types & Vectors)
**Video tutorial here**

## Topic 4.1: Data Types
R allows you to process and work with different types of data having different 
values. These are important types for our purposes:

- **numbers** (aka as **numeric values**)
- **strings** (aka as **characters**)
- **logical values**
- **missing values**

### Numeric values
Fundamentally important type of values are numbers. (R differentiates between 
different subtypes of numbers such as *double* and *integer*, which we ignore 
here). We have worked with numeric values before, for example when creating the 
object `start_of_WWI`.

When you want to know of what type an object is, you can use the `class()` 
function to determine it. 

```{r numbers, exercise=TRUE, exercise.lines = 3}
class(start_of_WWI)
```

### Strings
Strings, or character data, are data that is stored as text (individual letters,
words, sentences etc.). For example, party labels or city names can be stored 
as named objects in R. Note in the code below that, unlike the numeric values,
one assigns textual data to an object with the help of quotation marks. 
Quotation marks is the syntax that makes it clear to R that what comes in them 
is textual information.

For example, let us save the name of the current UN Secretary General in R:

```{r strings, exercise=TRUE, exercise.lines = 3}
UN_Gen_Sec = "António Guterres"

UN_Gen_Sec
```

### Logical values
R can also hold objects that contain logical values, such as `TRUE` or `FALSE`.
In the code below, for example, I compare two numeric objects and save the
result as an object that contains a logical value. Notice that the results of 
comparative operations can be automatically saved as an object:

```{r logicals, exercise=TRUE, exercise.lines = 10}
parl_gov = 21
pres_gov = 1

# comparing the values and saving the answer
num_parl_or_pres = parl_gov < pres_gov

# seeing the answer object
num_parl_or_pres
```


### Missing values
Finally, R has a special way of saving missing values under the label `NA`. 
`NA` stands for *not available*. For instance, in the code below we assign a 
missing value to an object about the elections in North Korea.

```{r missing_values, exercise=TRUE, exercise.lines = 3}
elections_North_Korea = NA

elections_North_Korea
```



## Topic 4.2: Vectors
Following the linguistic analogy, R also has nouns. Nouns are formed with 
letters. There are, unlike in a human language, only 4 types of nouns: 
**vectors**, **matrices**, **dataframes** and **lists**.

In this chapter, we look at vectors and matrices. 

### What is a vector?
A vector is a named object that can contain multiple values of the same type.
Because there are 4 different values (or letters), there are four types of
vectors: **string**, **numeric**, **logical** and **missing value** vectors.

Vectors in R are created by using a function called `c()`. `c()` stands for 
*concatenate* and does exactly that: concatenates several values into a single
named object. For example, in the code below we concatenate names of countries
in Africa that have elections in the first quarter of 2021 into an object 
called elections_africa:

```{r string_vector, exercise=TRUE, exercise.lines = 5}
# Creating a string vector
elections_africa = c("Uganda", "Somalia", "Central African Republic", "Nigeria", "Ivory Coast", "Republic of Congo")

elections_africa
```


Similar to the vector of strings, we easily create a numeric vector with the 
help of the `c()` function. In the example below, for instance, we create a 
vector called elections_months that contains the months in which the elections 
from the exercise above took place. Number `1`indicates that elections took 
place in January, `2` in February and `3` in March.

```{r numeric_vector, exercise=TRUE, exercise.lines = 5}
# Creating a numeric vector where
elections_months = c(1, 2, 2, 2, 3, 3)

elections_months
```



Lastly, using the same function we can create logical vectors. For example, 
let us answer whether the elections in the elections_africa were general, or 
not. `TRUE` indicates that elections were general while `FALSE` means that 
elections were not general:  

```{r logical_vector, exercise=TRUE, exercise.lines = 5}
# Creating a numeric vector
general_elections = c(TRUE, FALSE, TRUE, TRUE, FALSE, FALSE)

general_elections
```

**Naming Vectors**
One can also name vectors in R. When naming a vector, we assign a name to each 
element of a vector. It can be done with the help of a function called names. 
In the following example, we use the elections_vector as a way of naming the 
presidential_elections vector as follows:

```{r naming_vectors, exercise=TRUE, exercise.lines = 11}
# Vector of countries that held elections
elections_africa = c("Uganda", "Somalia", "Central African Republic", "Nigeria",
                     "Ivory Coast", "Republic of Congo")

# Vector indicating if the elections were general or not
general_elections = c(TRUE, FALSE, TRUE, TRUE, FALSE, FALSE)

# Naming the general elections vector
names(general_elections) = elections_africa
general_elections
```

As you can see in the printout, the vector general_elections has now a country 
name for each of the values in the vector.


### Vector selection


Sometimes we are only interested in some values that are contained in vectors. 
This requires that one learns to select a single or multiple values from a 
vector. There are two types of value selection in R. They are called 
*selection by position* and *selection by name*. 

When using the approach of selecting by position, we use a numeric index to 
indicate the value we are interested in selecting. To do so, we place the name 
of a vector object followed by the square brackets. And inside the square 
bracket, we locate the index. Using the general_elections vector from the 
previous exercise, let's select the first value in the vector:

```{r selecting_vector_position, exercise=TRUE, exercise.lines = 17}
# Vector of countries that held elections
elections_africa = c("Uganda", "Somalia", "Central African Republic", 
                     "Nigeria", "Ivory Coast", "Republic of Congo")

# Vector indicating if the elections were general or not
general_elections = c(TRUE, FALSE, TRUE, TRUE, FALSE, FALSE)

# Naming the general elections vector
names(general_elections) = elections_africa

# Selecting the first value in a vector
general_elections[1]

# Selecting the third value in a vector
general_elections[3]
```

Another approach to selecting a value from a vector is using a name. Similar to 
he example above, we use square brackets. But instead of using a numeric index, 
we indicate the name of a vector value we want to select. Pay attention to the 
fact that since names are string values, we type them in quotation marks. For
example, let's see if the elections in Uganda were general elections or not:

```{r selecting_vector_name, exercise=TRUE, exercise.lines = 17}
# Vector of countries that held elections
elections_africa = c("Uganda", "Somalia", "Central African Republic", "Nigeria", "Ivory Coast", "Republic of Congo")

# Vector indicating if the elections were general or not
general_elections = c(TRUE, FALSE, TRUE, TRUE, FALSE, FALSE)

# Naming the general elections vector
names(general_elections) = elections_africa

# Selecting the value for Uganda
general_elections["Uganda"]

# Selecting the value for Nigeria
general_elections["Nigeria"]
```



## Topic 5: Grammar of R (Matrices & dataframes)
Using only vectors to run complex analysis is inefficient. Thus, most of your 
coursework will involve collection of vectors on the same topic. R stores 
collections of vectors in several specific forms: matrices or dataframes. 
Matrices and dataframes are very similar. We will focus, however, on dataframes
in this chapter.

**Video Tutorial maybe** 


## Topic 5.1: dataframes

**1. Definition**  
A dataframe is the most common way of storing information for analysis.
Strictly speaking, it is a collection of vectors of equal length. Each vector 
in a dataframe usually represents a variable. The length of the vector then is 
the number of rows, or observations. A dataframe can consist of vectors
representing different data types: numeric, character, factor, logical. Data 
frames are essential for research. As an example, we will be working with a data 
frame that has information on parliamentary elections in Europe for the past 20 
years. The

1.1. Exercise: What is the difference between a matrix and a dataframe?
```{r quiz_matrix_dataframes, echo=FALSE}
quiz(
  question("What is the difference between the matrix and a dataframe?",
    answer("there is no difference"),
    answer("matrices hold less information than dataframes"),
    answer("dataframes can contain different types of information and matrices cannot", correct = TRUE),
    answer("matrices are not useful for data analysis")
  ))
```



**2. Creating a dataframe**

To understand dataframes better, it is helpful to create one by yourselves. 
As mentioned above, dataframes are created from vectors. I have prepared
several vectors for you to work with: *country_name*, *election_type*,  
*election_date*, *vote_share*, *party_name*, *left_right*, *seats*. As for
everything, there a function that helps to put vectors into a dataframe. It is 
called `data.frame()`. Its arguments, as you might have guessed, are vectors. 
Let's have a look at the function in more detail:

```{r data_frame_help, exercise=TRUE, exercise.lines = 4}
# find details of the data.frame() function
help(data.frame)  

```

As you can see in the arguments section, there are several useful arguments 
besides the component arguments. One of the most important ones is *row.names*. 
This argument allows one to specify whether one needs to use one of the columns
as row names or not. 

```{r creating_data_frames, exercise=TRUE, exercise.lines = 25}
# Load the individual vectors first
data("country_name")
data("election_type")
data("election_date")
data("vote_share")
data("party_name")
data("left_right")
data("seats")


# Creating a dataframes called elections_eu
elections_eu <- data.frame(country_name, election_type,
                           election_date, vote_share,
                           party_name, left_right, 
                           seats, 
                           row.names = NULL)

head(elections_eu)
```


**3. Looking at the dimensions of a dataframe**

After creating a dataframe, or importing it from another source, it is 
important to understand how it looks. There are several functions that can help 
with it. First of all, one can see how many rows and columns does a dataframe 
have with `dim()`, `nrow()`, `ncol()` functions. One can also look up the names 
of both rows and columns of a dataframe with the help of `rownames()` and 
`colnames()`. For example, let's investigate the cabinet dataframe: 

```{r df_dims, exercise=TRUE, exercise.lines = 13}
# loading the dataframe
cabinet <- data("cabinet")
head(cabinet)

# Dimensions of the data
dim(cabinet)
nrow(cabinet)
ncol(cabinet)

# Names 
colnames(cabinet)

```

Let's look at the row names separately, as the output of this function is going
to be long.

```{r df_rownames, exercise=TRUE, exercise.lines = 3}
# Names of columns
rownames(cabinet)
```



Find the dimensions of the dataframe you have created in the space below. And 
answer the questions that follow.

```{r df_dims_task, exercise=TRUE, exercise.lines = 5}
# Dimensions

# Names

```

**Assignment: dimensions of a dataframe**
```{r quiz_df_dims, echo=FALSE}
quiz(
  question("How many Rows and Columns are in the elections_eu dataframe?",
    answer("8648 columns and 7 rows", correct = FALSE),
    answer("86 columns and 48 rows", correct = FALSE),
    answer("7 columns and 8648 rows", correct = TRUE),
    answer("8646 obserations and 7 rows", correct = FALSE)
  ))
```



Understanding dimensions, however, is usually not enough. To decide on the type 
of analysis one wants to conduct on the data at hands, it is important to 
understand the types of data that are contained in a dataframe. To see more
detailed information on the variables in a dataframe, we can use the `str()` 
function. This function displays the structure of the dataframe with variables,
their types as well as a brief view of those variables: 

```{r df_str, exercise=TRUE, exercise.lines = 3}
# Investigating the structure of the dataframes
str(cabinet)

```

Now, investigate the structure of the elections_eu dataframe in the space 
below and answer the following question:

```{r df_stud_str, exercise=TRUE, exercise.lines = 3}
# Investigating the structure of elections_eu

```


**Assignment: dimensions of a dataframe**
```{r quiz_df_str, echo=FALSE}
quiz(
  question("Which of the following conclusions about your data is correct?",
    answer("There are 3 string variables, 3 numeric variables and 1 variable represening dates", correct = TRUE),
    answer("The dataframe contains numeric and string variables", correct = FALSE),
    answer("There are 4 numeric and 3 string variables", correct = FALSE),
    answer("There are 7 observations of 8648 variables", correct = FALSE)
  ))
```



Additionally, we can also just look at the first and last several observations 
of a dataframe with the help of two functions called `head()` and `tail()`:

```{r df_ht, exercise=TRUE, exercise.lines = 8}

# Investigating the top and bottom observations
head(cabinet)
tail(cabinet)

# Summary of the data
summary(cabinet)
```



Lastly, one can view even more detailed information about each variable in a
dataframe with the help of the `summary()` function. This function provides 
summary statisitcs about each variable in a dataframe. For example:

```{r df_summary, exercise=TRUE, exercise.lines = 3}
# Summary of the data
summary(cabinet)

```




**4. Rows and columns in a dataframe: selecting, replacing, renaming and adding values**

When working with dataframes, one often needs to select specific values of a 
dataframe. The logic and syntax of selecting elements from dataframes are
similar to the logic we used with vectors. That is, one can use square brackets
to select single or several elements from a dataframe. And most importantly,
just as with vectors we can use  *selection by position* and *selection by name* 
methods. There is one exception though.

Since dataframes are multidimensional - aka, they consist of rows and columns
- we have to account for that. The convention to separate rows from columns is 
to use a comma. The following grammar is fundamentally important to both 
understand and memorize `df[row, column]`. As you can see after specifying our 
dataframe, we use the space before comma to specify rows and the space after 
the comma to specify columns. Let's see how that works with one of our dataset.
We first select the second column and first row in their entirety. Afterwards, 
we look at the value in the first row and second column of the data:

```{r df_rc1, exercise=TRUE, exercise.lines = 13}
# loading the dataframe
cabinet <- data("cabinet")

# selecting the first row
cabinet[1]

# selecting the second column
cabinet[, 2]

# selecting the value in the first row and second column
cabinet[1, 2]

```


Now test your  understanding in the following snippet. Select the value which is 
located in the third columns and the tenth row of the cabinet dataset: 

```{r df_rc_e1, exercise=TRUE, exercise.lines = 4}


```

Columns in most of data-related research play a particularly important role. 
They represent the variables one is investigating. And that is why, they can be
also select by using he *selection by name* method. And as always, when using 
names - or string data for elements of a dataframe - they have to be specified 
within quotation marks:

```{r df_rc2, exercise=TRUE, exercise.lines = 4}
# selecting the first column
cabinet["country_name"]
```


Using *selection by name*, select the column from the dataframe which specifies
the number of seats in the cabinet dataframe:

```{r df_rc_e2, exercise=TRUE, exercise.lines = 4}


```



**5. Selecting multiple values at a time**

Most of the times when conducting an analysis, we are interested in looking at 
several values. We can also specify for selection of several values of dataframes 
in a similar manner to the selection of single values. To do that, one needs to 
specify in a vector form numeric values for rows and columns we are interested 
in looking up. For instance, in the snippet below, we want to see the first and
seventh rows as well as the third and fourth columns. We first select rows and
columns separately. In the last stage, we select the specified rows and columns
together. Pay attention to the fact that when we specify columns, R delivers the
specified columns while returning all rows. Similar logic applies to the
selection of rows. Selected rows are delivered together with all columns. That
is, if we do not specify a column or a row, R assumes that we want to see every 
row, or every column.

```{r df_rc3, exercise=TRUE, exercise.lines = 13}
data(cabinet)
# selecting the first and seventh rows
cabinet[c(1, 7)]

# selecting third and fourth columns
cabinet[, c(3, 4)]

# selecting the values which are located in rows 1 and 7, and columns 3 and 4
cabinet[c(1, 7), c(3, 4)]

```


In the following snippet, select (1) 8th and 9th rows, (2) 13th and 14th 
columns, and (3) then select the specified rows and columns from the previous 
steps at once:
```{r df_rc_e3, exercise=TRUE, exercise.lines = 10}


```



In the section above, we learned how to select two rows or columns at a time. 
Sometimes, one needs to select several values that are sequential. There is a 
special trick in R when it comes to selecting a sequence of values. For example, 
imagine that we want to look up all the columns starting from the 3rd till 11th.
To do that, instead of listing all of the columns, we use semicolons to tell R
that we want to select all the columns from a particular starting point to a 
particular ending point. Let's do exactly that with our data. In the example
below, we first select all columns between the 3rd and 11th columns as well as 
all rows between the 1st and 15th rows. Afterwards, we combine the row and 
column selection together. Pay attention to the fact that both the starting and
ending values are included in the resulting selection data. 

```{r df_rc4, exercise=TRUE, exercise.lines = 10}
# selecting columns starting from 3 to 11
cabinet[, 3:11]

# selecting rows from 1 till 15
cabinet[1:15, ]

# selecting rows 1:15 and col 3:11 at once

cabinet[1:15, 3:11]
```

In the following space, select (1) columns starting from 12th till 19th, (2)
rows starting from 1000th to 1050th and (3) values that correspond to the 
specifications in steps 1 and 2 at once. Also feel free to play around with
additional code

```{r df_rc_e4, exercise=TRUE, exercise.lines = 11}


```



An intermission quiz. We have a dataframe called `data`. Which of the
selection grammar correct?

```{r quiz_intermission, echo=FALSE}
quiz(
  question("Which of the following selection methods is correct?",
    answer("data['row number', 'column number']"),
    answer("data['column number', 'row number']"),
    answer("data[row number, column number]", correct = TRUE),
    answer("data['column name', 'row name']"),
    answer("data['row name', 'column name']", correct = TRUE),
    answer("data[data$column, data$row]")
  ))
```


**6. Adding information to a dataframe**

Once we can select values in a dataframes, adding information becomes easy. 
Imagine, for example, that the number of seats for Australian elections in the 
first 5 rows was recorded incorrectly. All the 5 first rows for that column 
should have a value of 13. To replace the values, we need to combine all the
elements of value selection we learned above into a single line of code. 
First, we specify the dataframe's name, then the column we are interested in 
and then we select the sequence of values and then assign the proper value to
the selection. Here is how we can replace it:


```{r df_ai1, exercise=TRUE, exercise.lines = 4}
# replacing a sequence of values in a column
cabinet$seats[c(1:5)] <- 13

```

Now, let's imagine that the values for 1st and 5th rows of the seats column 
needs to be replaced by 10. We can use the same logic as above to do so:

```{r df_ai2, exercise=TRUE, exercise.lines = 4}
#replacing the 1st and 5th row values
cabinet$seats[c(1,5)] <- 10

```

In the following snippet, (1) replace the values in rows 9-15 of the column s
eats with a number 14 and (2) replace the first five values of the 
`election_seats_total` column with a number 80

```{r df_ai_e1, exercise=TRUE, exercise.lines = 11}


```


**7. Adding Columns and rows, as well as renaming them**

Let's start with renaming. There are special functions that help to name both 
columns and rows: `colnames()`, `row.names()`. For both functions, we have to 
specify the dataframe and the column number, and then assign a new name. For
example, let's rename `election_seats_total` which is the 10th column to 
`total_seats`:

```{r df_crn2, exercise=TRUE, exercise.lines = 4}
# rename the column 
colnames(cabinet)[10] <- "total_seats"
head(cabinet)

```


Changing the names of rows follows the exact same logic as above, only with its
own function:

```{r df_crn3, exercise=TRUE, exercise.lines = 4}
# rename the row 
row.names(cabinet)[1] <- "First Row"
head(cabinet)

```


In the space provided below, rename the column `country_name` to 
`country_name_long`. Additionally, rename the 2nd row to the `Second Row`:

```{r df_crn4, exercise=TRUE, exercise.lines = 4}
# rename the column

# rename the row


```


Another intermission quiz. Which of the functions below is not a function?

```{r quiz_intermission1, echo=FALSE}
quiz(
  question("Which of the following is not a function?",
    answer("rownames()"),
    answer("row.names()", correct = TRUE),
    answer("colnames()", correct = TRUE),
    answer("col.names()"),
    answer("columnnames()")
  ))
```



**8. Adding entire columns and rows to a dataframe.** 

Adding columns to existing dataframe is easy if one pays attention to one 
detail: the length of the vector that we are trying to add as a new column 
should be equal to the length of the dataframe. There are two ways of adding 
a column. We explore both in the snippet below. 

Our cabinet dataframe's length is 12085. That is there are 12085 row values for
each variable, or column. Let's add a vector of that length as a column to our 
dataframe. I will create a simple numeric vector of length 12085 for the sake
of an example. 

The same vector will then be added first by using the assignment operator. And 
then using a special function called `cbind()`. `cbind()` stands for column 
bind. `cbind()` does exactly what it says: it binds two objects by column. In 
the example below, we are binding a vector `fictional_vector` to our `cabinet` 
dataframe. We could have used `cbind()` to bind an entire new dataframe to
the `cabinet` (*again paying attention to the fact that both of the dataframes
should have then the same length, or number of rows*).

```{r df_crn5, exercise=TRUE, exercise.lines = 12}
data(cabinet)

# create a numeric column
fictional_vector <- c(1:12085)

# add the column to the dataframe
cabinet$added_vector <- fictional_vector

# functional way of adding a column
cabinet <- cbind(cabinet, fictional_vector)

```


We have prepared a vector called `party_name`. The vector provides an augmented 
party names for the `cabinet` dataframe. Add that vector to the `cabinet` 
dataframe. You can choose to add either with the help of the assignment operator
or the `cbind()` function. 

```{r df_crn6, exercise=TRUE, exercise.lines = 3}
# add a new column to the dataframe


```



Similarly, rows can be added to a dataframe. Usually, large dataframes contain 
a lot of columns that adding row information becomes tedious with the assignment 
operator method. Thus, it is recommended to use the `rbind()` function to add 
rows. This function is a twin of the `cbind()` but instead of binding data 
objects by columns, it binds them by row. Once again, the vector we are adding 
as a row should contain values for each of the corresponding columns. Our 
dataframe, as you remember, contains 18 columns. So, the length of the new row 
vector we add should also equal to 18. Let's create and add such a vector.

In the example below, I am adding a row for the last Australian elections that 
were not included in the original dataframe. The vector containing the 
information is called `AUS_last`. Feel free to explore the vector, it is 
available in your working environment. In the example below, I explicitly check 
the length of the vector to make sure that I am capable of combining it to the 
`cabinet` data. As mentioned above, if the lengths differ, adding new 
information is next to impossible.

```{r df_crn7, eval = F, exercise=TRUE, exercise.lines = 11}
# checking the length of the AUS_last vector
length(AUS_last) 

# adding the row to the cabinet dataframe
rbind(cabinet, AUS_last)
```

It turns out, the fist elections for Bulgaria were also not added to the 
`cabinet` dataframe. We have added the vector for the first Bulgarian elections 
in your working environment under the name `BG_first`. In the following snippet,
add it to the `cabinet` data. 

```{r df_crn8, exercise=TRUE, exercise.lines = 11}

# incorporate the BG_first vector into the cabinet dataframe

```



## Topic 6: Packages

Packages are akin to books that are specific to a field. Packages consist of 
functions(verbs) that perform specific analysis. For instance, there are
packages that have functions which allow one to conduct a regression analysis, 
visualize data, organize and clean data, conduct other types of analysis. These
field-specific books are available via an R-specific library called 
[CRAN](https://cran.r-project.org/). 

When using packages in R, one has to remember two important things. First of
all, one has to download packages from CRAN. This is done with a special 
function `install.packages("package name")`. As you can see in the example below,
one has to make sure to incorporate the name of the package in quotation marks:

```{r pack_ins1, eval = F, exercise=TRUE, exercise.lines = 7}
# example of installing a package
install.packages("ggplot2") # ggplot is a package that allows one to visualize data

# installing the ggparliament package
install.packages("ggparliament")

```

The second important point about packages is the need to explicitly open them, 
as one would open a book in order to be capable of reading it. To open a 
specific package in R, one, of course, uses a function called 
`library(package name)`. Here, unlike, when downloading packages from CRAN, 
there is no need for quotation marks. Once again, every time we want to use a 
specific package that we already downloaded earlier, one has to open the book 
before using it. In the example below, we create a simple frequency graph of 
our cabinet data with a visualization package `ggplot2`:

```{r pack_ins2, eval = F, exercise=TRUE, exercise.lines = 7}
# opening the ggplot2 package
library(ggplot2)

# opening the ggparliament package
library(ggparliament)

```

As mentioned in the chapter introduction, packages are like books that contain 
functions. Those functions allow us to work with data on many levels. The two
packages we downloaded and opened in the sections above specialize in
visualization. Now, let's use the `ggparliament` package. In addition to special
functions for graphing parliamentary data, this package also incorporates a 
dataframe called `election_data`. In the snippet below, we use both some of that
data and functions from `ggplot2` to plot the parliamentary composition of 
Germany in 2017 in the form of a "horseshoe" graph:

The following code is an example of what packages can do. 
**Please, do not try to memorize the code, there is a special module on
visualization within your program. That module covers the details of `ggplot2` 
in a more applied manner!**

```{r pack_ins3, eval = F, exercise=TRUE, exercise.lines = 31}
# Opening the books
library(ggparliament)
library(ggplot2)

# Selecting the data for German Bundestag for 2017
DE <- election_data %>%
    filter(country == "Germany" &
               house == "Bundestag" &
               year == 2017) 

# Reforming the data to fit the type of a plot
DE_horseshoe <- parliament_data(election_data = DE,
                                       party_seats = DE$seats,
                                       parl_rows = 8,
                                       type = "horseshoe")

# Plotting the data
DE_plot <- ggplot(DE_horseshoe, aes(x, y, colour = party_short)) +
    geom_parliament_seats() + 
    theme_ggparliament() +
    labs(colour = NULL, 
         title = "Bundestag") +
    scale_colour_manual(values = DE$colour, 
                        limits = DE$party_short) + 
    theme(legend.position = 'bottom')

# Printing the plot
DE_plot
```


In the space below, install two packages called `dplyr` and `magrittr`. We will
study some contents of both packages in the upcoming chapter of this module. 


```{r pack_e, exercise=TRUE, exercise.lines = 3}
# install the packages


```



## Topic 7: Tidy Data & Data manipulation

### Introducing dplyr

This chapter is about a special package called `dplyr`. It is a package that 
specializes in helping one to manipulate data (subset, select, filter and many
more functions). The package provides functions that make working with dataframes
easier and more intuitive as compared to the basic R approach (which we have 
covered in the chapter on dataframes). On the official 
[website](https://dplyr.tidyverse.org/) of the package, it is referred to as a
package that provides "a consistent set of verbs that help you solve the most 
common data manipulation challenges". 

In this chapter of the module, we are going to cover five most important 
functions within the dplyr package. These functions will be useful for all of 
the modules available to you in this program and data manipulation within R in
general. 

The last task of the previous chapter was installation and loading of dplyr. 
If you have not done so let's start by installing and loading the package:

```{r dplyr1, eval = F,  exercise=TRUE, exercise.lines = 7}
# install the package
install.packages("dplyr")

# load the package 
library(dplyr)
```


### Five main functions of dplyr: select(), filter(), arrange(), mutate(), summarize()

**2.1: select() for columns**

First of the functions we cover is `select()`. Select is a function that allows 
one to select a variable from a dataframe. This is an alternative to the 
selection we learned in the chapter on dataframe manipulation. The function 
allows to select one as well as multiple variables at a time. Do not forget that
variable is a column of a dataframe. 

The `select()` function has two important arguments: a dataframe and the name, 
or names, of columns we are interested in selecting. In the example below, let's 
select `seats` variable from the `cabinet` dataframe. We can specify that 
`cabinet` is our dataframe by explicitly assigning it to the `.data` argument 
within the `select()` function. It is not a must, however. `Dplyr` knows that 
the first object inside the brackets is a dataframe on its own (see the second 
example in the code snippet). Pay attention to the fact that we do not have to 
use the quotation marks when selecting a column with `dplyr` functions:


```{r dplyr2, exercise=TRUE, exercise.lines = 5}
# selecting a single variable
select(.data = cabinet, seats)

# selecting without specifying the data explicitly
select(cabinet, seats)
```

Now, we can also select multiple variables (or columns) at once with the help
of select function. In addition to the seats variable, let's select the variable 
called `election_date` and assign the selection to a new dataframe called 
`cabinet_subset`:


```{r dplyr3, exercise=TRUE, exercise.lines = 3}
# selecting a single variable
cabinet_subset <- select(.data = cabinet, seats, election_date)

```

**Exercise**
In the space below, exercise your knowledge of the `select()` function by 
selecting the variables called `seats`, `election_date`, and `country_name`.
Do not forget to assign the selection to a new dataframe called `cabinet_sub`.
We will use the created subset in the end of the chapter to create a chart. 

```{r dplyr4, exercise=TRUE, exercise.lines = 3}
# selecting and assign


```


**2.2: filter() for rows**


Now, `select()` is very useful for working with columns or variables of a 
dataframe. As you might remember from the earlier chapters, however, one 
sometimes needs to subset dataframes based on the row values. `Dplyr` has a 
special function that allows one to easier subset the data based on the row 
values as well. And it is called `filter()`. 

**Filtering for numeric columns**
Similar to `select()`, the first argument for `filter()` is the dataframe.
The consequent argument, however, is the name of a variable followed by the row 
value we are interested in selecting. As an example, let's create a subset of 
the `cabinet` so that it only includes all the data but for a single election
year: 2017. Date information is stored in the column called `year`. So, let's 
filter for the value `2017` in the column `year`, and save the subset in a new 
dataframe called `cabinets_2017`: 

```{r dplyr5, exercise=TRUE, exercise.lines = 5}
# filtering for Germany
cabinets_2017 <- filter(cabinet, year == 2017)

```

**Filtering for string columns**
As you might have noticed, the grammar (arguments) of `filter()` is quiet s
imilar to `select()`. It becomes a bit different, however, when we want to 
subset the data based on the values of a string variable. Namely, when selecting
text data, the values we are interested in filtering on needs to be included in 
quotation marks. Let's, see an applied example by subsetting the cabinet
dataframe first for year `2017`. In addition to selecting the year, let's also 
filter for a single country: Germany. Country information is stored in the 
column `country_name`, and the value we are interested in is `Germany`. We will
save the subsetted data in a new dataframe called `cabinet_DE_2017`:

```{r dplyr6, exercise=TRUE, exercise.lines = 5}
# filtering for Germany
cabinet_DE_2017 <- filter(cabinet, year == 2017, country_name == "Germany")

```


**Filter assignment**
To practice your filtering skills, create a new dataframe in the chunk below 
called `cabinet_AUS_2001` which includes the subset of the `cabinet` data for 
Australian cabinet of 2001.

```{r dplyr7, exercise=TRUE, exercise.lines = 3}
# filtering Australia in 2001

```


**2.3: arrange()**


One of the most simple functions in `dplyr` package is the one called arrange.
It is a function that becomes extremely useful for cases when the data one works 
with comes in an unordered manner. `arrange()` as the name suggests let's the
researcher to arrange a dataframe. The function allows to arrange an entire 
dataframe based on the values of a single column. And the direction of the
arrangement can be specified with an additional argument that allows us to
indicate wheter we want to arrange the data in ascending or descending order.

As an example, let's arrange the cabinet dataframe based on the values of seats.
In the example below we will (1) first subset the data such that only 
`country_name`, `year`, `party_name_english` and `seats` variables are *selected*
and (2) secondly we will *filter* the data to only include values for Bulgaria. 
Let's name the subset data BG_subset. 


```{r dplyr8, exercise=TRUE, exercise.lines = 7}
BG_subset <- filter(cabinet,country_name == "Bulgaria")
BG_subset <- select(BG_subset,country_name, year, party_name_english, seats)


```


As you can see in the subset, the data seems to be ordered by the year of 
election. This data can be easily reordered based on the values of the
`party_name` variable with the `arrange()` function as follows:

```{r dplyr9, exercise=TRUE, exercise.lines = 3}
# arranging by party name
arrange(BG_subset, party_name_english)
```

Re-arranged data now is ordered by the party name. Since we are arranging by the
string variable, the data is arranged alphabetically. If we want to look at the 
data in the reverse order, one only needs to add an internal function of the 
`arrange()` called `desc()`. One needs to simply include the name of the variable
by which we are arranging in the `desc()` function. It is important to remember
that `desc()` is an internal funciton of the `dplyr` function. *Thus, it will
not work in other contexts.* Once again, the `desc()` as its name suggests, 
reverses the arranging in the descending order. Let's see the application with 
the previous example:


```{r dplyr10, exercise=TRUE, exercise.lines = 3}
# arranging by party name
arrange(BG_subset, desc(party_name_english))
```

Pay attention to the fact that while, the resulting data is arranged by the
name of the party, the year column is not arranged (that is there is no
descending or ascending order according to the year).


Finally, the beauty of `arrange()` is that it allows for ordering the data 
based on values of several columns. Imagine, for example, that we want to 
arrange our subset data based on two variables: party name and year. Once we 
run the code below, we can see that the data is first arranged by the party name 
alphabetically and then by year in reversed chronological order


```{r dplyr11, exercise=TRUE, exercise.lines = 7}
# arranging by party name and reverse chronologically
arrange(BG_subset, party_name_english, year)

# the same arrangement, except the year is now chronological
arrange(BG_subset, party_name_english, desc(year))
```

**Exercising `arrange()`

In the space below, filter the `cabinet` data for values of Germany only. Then
select the variables: `column_name_english`, `year`, and `seats`. Name the
subset `DE_cabinet` and arrange the data by the number of seats and party name.
The data should be arranged such that the parties with the highest number of 
seats are presented.

```{r dplyr12, exercise=TRUE, exercise.lines = 15}



```


**2.4: group_by() and summarize()**

Another function of `dplyr` helps to create a quick summary of the data 
available. The function is called `summarise()`. 

The penultimate function we cover in this section is the `summarise()` from
`dplyr`. This function is useful for deriving summary data for a dataframe. 
Usually, the function is used in combination with functions that provide summary 
statistics, such as the `mean()` for finding the average, `min()` or `max()`for 
finding the minimum or maximum, and others.  Similar to other `dplyr` functions,
summarize requires the naming of the dataset as a first argument. And as a 
second argument, it takes in a function which helps us calculate the summary 
statistics we are interested in together with a variable for which the summary 
statistic is required. Let's calculate the average number of seats in our 
`cabinet` data as an example: 


```{r dplyr13, exercise=TRUE, exercise.lines = 15}
# calculate the mean number of seats in our cabinet data
summarise(cabinet, mean(seats))

```

While this is useful, you might wonder why one needs to use summarize in the 
first place. Since, we could have calculated the mean for seats of the cabinet 
data in a simpler way by simply including the seats column inside the `mean()` 
function:


```{r dplyr14, exercise=TRUE, exercise.lines = 3}
# simpler way of finding the average

mean(cabinet$seats)

```

The contradiction is valid. The `summarise()` function's actual usefulness 
arises when we want to see summary statistics based on some sort of grouping.
Imagine, for example, that we want the same average number of seats but for each
country. If we followed the base R, we would have to first create a subset for 
each country of our dataframe and then use the `mean()` function with each of
the subsets. As a glimpse to how it would look like, let's create such subsets
for the first three countries (Australia, Austria and Belgium) and calculate the 
mean with base R:


```{r dplyr15, exercise=TRUE, exercise.lines = 20}
# Subsets
# subset for AUstralia
AU_subset <- filter(cabinet, country_name == "Australia")

# Subset for AusTria
AT_subset <- filter(cabinet, country_name == "Austria")

# Subset for Belgium
BE_subset <- filter(cabinet, country_name == "Belgium")

# Averages 

# for AU
mean(AU_subset$seats)
mean(AT_subset$seats)
mean(BE_subset$seats)

```

Given that our data contains 37 countries, the seemingly simple tasks becomes 
very repetitive which leads to boredom and errors. To calculate the means for 
each of the countries in this example, we would have to create 37 subsets and 
calculate 37 means (which would result in 74 lines of code). This is where
`summarise()` in combination with another `dplyr` function called `group_by()` 
becomes really useful. The ability to combine these two functions leads to 
substantial as well as temporal efficiency, as we only need to write a relatively
short piece of code. 

Before moving on to combining the two functions, let's first have a more 
detailed look at the `group_by()`. This is another function that is internal 
to `dplyr` similar to `desc()` that we have covered above. `group_by()` only 
makes sense within the `dplyr` environment and suggestively, it groups the data
based on a certain column. It is important to know that nothing in the dataframe 
changes when we apply the `group_by()` on its own. See below, for instance:

```{r dplyr16, exercise=TRUE, exercise.lines = 5}
# group_by introduction
group_by(cabinet, country_name)

```

The only thing that happens is that dplyr tells R to treat the `cabinet` data as 
grouped by `country_name` column internally. This internal functionality becomes 
useful when we want to see summary statistics for data that is grouped by values 
of that column. That is, let's group the `cabinet` data by `country_name` again.
And then calculate the average `seats` for the grouped data: 

```{r dplyr17, exercise=TRUE, exercise.lines = 7}
# group_by icountry
grouped_cabinet <- group_by(cabinet, country_name)

# calculate the mean for grouped data
summarise(grouped_cabinet, mean(seats))

```

As you can see, the last line of code, provides us with the mean number of seats 
for each country in our dataset. This is so because we grouped by country and
then calculated the mean. We have reduced the potential 74 lines of code to two 
lines.

Lastly, `group_by()` allows to group the data by several variables. This further 
leads to an opportunity to get even more detailed summary view of the data. As
an example, let's group the `cabinet` data not only by country but also by 
election year. And see, what summary view of the mean for seats would deliver:

```{r dplyr18, exercise=TRUE, exercise.lines = 7}
# grouping by country, year
grouped_cabinet <- group_by(cabinet, country_name, year)

# calculate the mean for grouped data
summarise(grouped_cabinet, mean(seats))

```


Because we first grouped the data by country and then by year, as you can see,
we could obtain the mean seats for each country for each year!

**group_by and summarize exercises** 
In the space below, calculate (1) the average seats occupied by each party 
within each country; (2) next calculate the average seats for each party per 
year. Save the results of the calculation in a new dataframe. You can choose 
the names for each of the resulting data yourself.


```{r dplyr19, exercise=TRUE, exercise.lines = 13}



```


**2.5: mutate() for creating new values**

Finally, we are covering the last function of the `dplyr` package within this
module. The function is called `mutate()`. It is the only function in the
package that allows one to create and simultaneously add a totally new variable 
to the dataframe we are working with. In other words, it allows one to mutate a 
dataframe a bit. 

As a case, let's calculate the proportion of seats gained by a single party in
an election year to the total number of seats in that year. To achieve the task,
we will first do some subsetting with the functions we have learned in the
previous step. First of all, let's only select the variables that are needed for 
the work: `country_name`, `party_name_english`, `year`, `seats`, and 
`election_seats_total`. Secondly, since we want to derive the data per party per
year, let's do the needed grouping of the data:


```{r dplyr20, exercise=TRUE, eval = F, exercise.lines = 23}
data("cabinet")

# Subset the data first
cabinet_subset <- select(cabinet, country_name, party_name_english, 
                         year, seats, election_seats_total)

# Group the data
cabinet_group <- group_by(cabinet_subset, party_name_english, year)

# Investigating the data
head(cabinet_group)

```

Notice that the subset and grouped data has 5 columns that we have selected 
and is grouped by party names and election year. 


Next to calculate the proportion, we employ the `mutate()` function. Similar 
to the rest of `dplyr` functions, the first argument of `mutate()` is the 
dataframe we are working on. The second argument of `mutate()` is the name of 
the variable we want to create. Since we are calculating the proportion of 
seats, we will call our variable `seat_proportion`. Following the name of the
new variable we want, we have to write the formula that we employ to calculate 
the value. To find the proportion of seats in our case, we divide the `seats` 
column by the `election_seats_total` column. Here is how it all looks in code:


```{r dplyr21, eval = F, exercise=TRUE, exercise.lines = 7}
# mutating the data
cabinet_mutated <- mutate(cabinet_group, seat_proportion = seats/election_seats_total)

```


**Assignment for the mutate** 

To exercise your skills at mutating a dataset, create two new variables based 
on the `seats` column of the `cabinet` dataset in the space provided below. The 
first variable, called `difference` should equal to the difference between the
mean seat number per country, per election year and the individual seat number 
gained by each party in that year. The second variable called `squared.diff` 
should equal to the square of the variable you have created in the previous step.

```{r dplyr22, exercise=TRUE, exercise.lines = 33}




```


Congratulations! If you made it until here, you have mastered foundational skills needed to work on data-related assignments. 

## Topic 7.1: Advanced tidy: the pipe operator
This section is meant to be a breathing space where we get acquainted with a 
new operator: **pipe operator**: `%>%`. The pipe operator is a special operator 
that belongs to the `magrittr` package. Pipe operator is one of the tools that 
makes it easy to write human-understandable and elegant code, especially when 
using functions that belong to `dplyr`. The only purpose of the pipe operator 
is to reduce the intermediate data creation during analysis. It achieves the 
reduction by automatically turning the input of one line of code into the input 
of the following piece of code. A linguistic proxy to the pipe operator is the 
adverb **then**.

Let's see how to use it in an applied manner. If you have successfully completed 
the last coding task, your output should have looked something similar to the
code below:


```{r pipe, eval = F, exercise=TRUE, exercise.lines = 33}
# loading the data
data(cabinet)

# selecting the needed columns
cabinet_subset <- select(cabinet, country_name, party_name, year, seats)

# grouping
cabinet_grouped <- group_by(cabinet_subset, country_name, year)

# mutating the difference between the seats and the average
cabinet_mutant_first <- mutate(cabinet_grouped, difference = seats - mean(seats))

# mutating the squared difference
cabinet_mutant <- mutate(cabinet_mutant_first, squared.diff = difference^2)

```

In this code, what we want to achieve is the final dataset called `cabinet_mutant` 
which has two new variables: `difference` and `squared.diff`. Pay attention to
the fact that each of the intermediary subsets of the data -- `cabinet_subset`,
`cabinet_grouped` and `cabinet_mutant_first` -- are all used as first arguments 
for consequent steps. But outside of that, they are needless. In fact, they can
also be harmful as we engage in creating a lot of subset with somewhat similar 
names. This can lead to errors, confusion and if you are working with very large 
datasets, your computer might slow down considerably. 

Pipe operator is useful exactly in these types of situations: it helps us to 
automatically use the results of one line of code as an input (or first argument)
of the following piece of code. Let's rewrite the code above with the pipe.

We first start by typing the name of the final dataframe we want: `cabinet_mutant`. 
Secondly, we start by applying the `select()` on the original `cabinet`data. And 
then we chain each of the consequent tasks - `group_by()` and `mutate()` - with 
the pipe operator as follows: 


```{r pipe1, eval = F, exercise=TRUE, exercise.lines = 33}
# loading the magrittr library
library(magrittr)  # %>% operator comes from magrittr

# selecting the needed columns
cabinet_mutant <- cabinet %>% # take the cabinet data, THEN
  select(country_name, party_name, year, seats) %>% # select the columns from the cabinet, THEN
  group_by(country_name, year) %>%                  # group the data, THEN
  mutate(difference = seats - mean(seats)) %>%       # calculate the difference with mutate, THEN
  mutate(squared.diff = difference^2)       # calculate the square of difference with mutate

# let's check the created data
head(cabinet_mutant, n = 25)
  
```


**Assignment for the pipe operator**

Using the pipe operator, create a subset of the `cabinet` dataframe such that 
(1) only value for France in 2012 are present for (2) the following variables: 
country_name, year, party_name_english, seats, and left_right. Name the subset 
dataframe `cabinet_FR`.

```{r pipe2, exercise=TRUE, exercise.lines = 23}
# Create the defined subset with the pipe operator

  
```

## Topic 8: Import data into R
Up to this point in this module, you have used datasets provided by us directly 
to your working environment. When working with data tasks for your classes, 
however, you will have to import various datasets. *Importing a dataset* means 
opening a file that contains data wit the help of R. 

As you might have guessed, there is a special book that helps one import data in
R. In fact, there are two important packages called `readr` and `readxl`. As all 
the other packages, you can install and *have to load them* as we do in the 
following snippet. Importantly, the following code has both packages loaded, 
usually you would not need them both. 

```{r readr, eval = F, exercise=TRUE, exercise.lines = 10}
# Installing and loading the readr package
install.packages("readr")
library(readr)

# Installing and loading the readxl
install.packages("readxl")
library(readxl)
  
```

Now, both of these packages have multiple functions that let one import datasets
in R. If you are wondering why is there a need for two different packages and a 
bunch of functions at that, you are at the right point. The fact is that datasets
are stored in different formats. All of you have probably seen datasets in 
Excel files. It is one way of storing rectangular data. Excel is not the only 
format, however. In fact, there is a plethora of formats, such as
**.csv (comma-separated value file)**, **.rds(specific file in which R stores data)**, 
**.txt(text value file)**, **.xls or .xlsx (excel files)** as well as 
**.dta (files specific to stata data)**. One can think of those formats as sort
of family names. Collectively, these *family names* are referred to as extensions.
As an example, if you have received the cabinet data in an excel format, the 
file would probably have such a name: **cabinet.xlxs**. 

Extensions are important because they signal to different software, such as R,
Stata or even Microsoft Excel what type of data format they are dealing 
with. Without such specification, one can ruin the dataset when opening it. 
That is the reason, for instance, why opening `.csv` files in Microsoft excel 
leads to problems (especially if you have the European-licensed Microsoft office). 
**As a rule, do not open your csv files with excel**.

These extensions are also important in R environment. Unlike most commercial 
software, R can read all of the data formats correctly if one uses the correct
package and functions. That is the reason why there are different packages for
reading in data, such as `readr` and `readxl`. Now, we will cover two important 
functions from `readr` and from `readxl`. Most of data you will be working with 
will come either in the **.csv** or **.xls** formats. 

Let's start with the **.csv** and `readr`. The most important function from 
`readr` is called `read_csv()`. Intuitively, it reads in csv files. We have
placed a dataframe saved in the **.csv** format, it is called 
**country_name.csv**, let's see how to read it. When reading in a dataset, we 
first to have create a name with which we want to use it. Let'S call ours 
`country_name`. Then we assign to that name the output of a function `read_csv()`.
The first argument inside the function is the name of the data file as it is
stored in a folder on our computer, or on any other source. Once again, since
we are dealing with names, we have to put the name of the data file in quotation 
marks:

```{r readr1, eval = F, exercise=TRUE, exercise.lines = 7}
# Reading in the country_name.csv
country_name <- read_csv("country_name.csv")

# inspect the data
head(country_name)
  
```

Congratulations, you have just **imported** your first data! Now, `read_csv()` 
has additional useful arguments. We will cover two of the most important, called
`col_names` and `skip`. The first of these, `col_names`, accepts as a value either 
`TRUE` or `FALSE`. By default (automatically), the value is set to `TRUE`, and as
such `read_csv()` basically treats the first row of the original dataset as 
column names. Sometimes, however, the first row of a dataset is not a column 
name, we sometimes get the data where column names are not included. In such
cases, we can set the `col_names` argument to `FALSE` as in the example below:

```{r readr2, eval = F,  exercise=TRUE, exercise.lines = 10}
# Reading in by changing the col names
country_name2 <- read_csv("country_name.csv", 
                         col_names = FALSE) # we told R not to treat
                                            # the first row as col names

# Now look at the again
head(country_name2)
  
```

As you might have noticed upon the inspection, the data does not have column 
names anymore, and the first row - country_names - is treated as a value. 


## Topic 9: Updating R & RStudio

Updates of R and Rstudio are recommended to be done at least once a year. As for 
everything, there is a special package `installr` (for Windows users). The 
`installr` package as a function called `updateR()` which does exactly what it 
says: updates R to the most recently available version. If you were updating R 
on your own system, the code would look as follows. **Do not run the following
code now** but you are encouraged to play with it on your own system:

```{r installr, eval = F, exercise=TRUE, exercise.lines = 23}
# install the installr package
install.packages("installr")

# Always load the package
library(installr)

# Update R
updateR()

```

For Mac users, one can simply download the newest version of R from CRAN and the 
system will automatically shift to using it. If you do not remember how to 
install R, look at the first chapter of the tutorial.

Updating RStudio happens via the Help tab on the menu of RStudio both for Mac
and Windows users. 

**Note to the team and myself**: maybe it makes sense to include RStudio update 
in the video tutorial.

Congratulations, you are done with this module!

---
